---
title: "Comparing Kinases"
author: "Kieran Mace"
date: "December 2, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=11, fig.height=11, message = FALSE)
source('../utils/load_libraries.R')
source('../utils/load_functions.R')
library(d3heatmap)
library(IDPmisc)

t2g = load_transcripts_to_genes()

master_meta = get_sample_meta()
#raw_counts = load_count_matrix(master_meta,'star')

# lib_size = apply(raw_counts,2,sum)
# p = t(apply(raw_counts,1,function(x) x/lib_size))
# p = rename_gene_names(p, t2g)

```

## Extracting difference in Kinase within the Same Stress

The primary goal of this project is to find kinases that, when inhibited, change some aspect of gene expression.

These effects may be classified into two groups. those that happen independant of the type of stress, and those that only happen within a given stress.

But before we attempt to disect these differences, we must first attempt to disect differneces in general. 

I will therefore attempt to analyze a few stresses, starting with heatshock, to identify potential kinases that affect gene expression under this stress.

### Heatshock

We will first extract out all the data that is under heatshock, with the Cocktail. We can either:
1. look at this data raw, 
2. normalize by the WT strain, 
3. or normalize by each strain with just drug. 

lets start with Raw:

But lets only look at the genes with the highest CV^2^

```{r heatshock_raw, cache=TRUE}

interactive_heatmap = function(plate_code){
meta = master_meta %>% filter(Plate_Code==plate_code)
raw_counts = load_count_matrix(meta,'star')
dds = DESeqDataSetFromMatrix(countData = raw_counts, colData = meta, ~Strain)
dds = DESeq(dds)
rld = rlog(dds, blind=FALSE)
library(genefilter)
topVarGenes <- head(order(rowVars(assay(rld)),decreasing=TRUE),200)
mat <- assay(rld)#[ topVarGenes, ]
mat <- mat - rowMeans(mat)
mat = rename_gene_names(mat,t2g)
colnames(mat) = meta$Strain_Code
#df <- as.data.frame(colData(rld)[,c("Strain")])
#pheatmap(mat, annotation_col=df)
mat[mat > 2] = 2
mat[mat < -2] = -2
d3heatmap(mat, colors = "RdBu")
}

# Drug SDC
interactive_heatmap("d")
# Drug YPD
interactive_heatmap("yd")
# Drug Heatshock
interactive_heatmap("h")
# Drug Glucose Dropout
interactive_heatmap("g")
# Drug Azetidine
interactive_heatmap("a")
# Drug Fluconazole
interactive_heatmap("f")
# Drug Fluconazole
interactive_heatmap("m")

```



```{r heatshock_wt_norm}
```
## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
