---
title: "RNA Quality Check"
output:
  html_document:
    toc: true
    fig_caption: true
---

## Overview
One thing we need to look into is how reproducible the data is.

In order to do this, I will look at data that concerns Menadione.

## Analysis
First I must load the data:

```{r data_loading, message=FALSE, warning=FALSE}
rm(list=ls())
source('../utils/load_libraries.R')
source('../utils/load_functions.R')

t2g = load_transcripts_to_genes()

master_meta = get_sample_meta()
raw_counts = load_count_matrix(master_meta,'star')

lib_size = apply(raw_counts,2,sum)
p = t(apply(raw_counts,1,function(x) x/lib_size))
```

### Looking at WT Strains
Under WT, I have three types on Menadione data.

WT in Tubes done by Xu
WT in Plates with Drug Cocktail done by Kieran
WT in Plates done by Kieran

First question is do these data cluster across condition.

I will first look at the raw expression of these. disregarding any baseline normalization:

For Reference:
WT_MEN is done by Xu, in tubes - Dose unknown
m is done by Kieran, in plates, with Drug
k is done by Kieran, in plates, without Drug

```{r scatter_1, fig.width=11, fig.height=11, cache=TRUE}
check_stress = function(stress) {
  meta = master_meta %>% filter(Strain=='WT', Stress==stress)
  pairs(log2(p[,meta$Sample_Name]),pch=20)
}
check_stress('Menadione')
```
Here we see that the differences within group are very small. meaning that the source of differences between these groups are due to the experimental proceedure, which include: day of experiment, drug inclusion, and experimenter.

One thing I notice, however is that the K set, are less similar to Xu's data then the M set (M set has the drug in it). This is not to say that Drug could be a replacement for shaking, but it does indicate that the increased stress caused by drug might reflect a similar increase in stress from shaking. 

To drive this point home, I will also look at data that I have for two other conditions: Azetidine, and Heatshock
```{r scatter_2, fig.width=11, fig.height=11, cache=TRUE}
check_stress('Azetidine')
check_stress('Heatshock')
```
One consideration here, is that there is no baseline in these plots, I am simply plotting the raw values.

There is one obvious problem with this. the axis of these plots have ranges of ~10 orders of log2 magnitude. so even large fluctuations (4 fold change) will only vary in this plot by 2 units. However it is important to stress that may comparrision of RNA-seq libraries look like this. quoting values of R^2^

One quick way to get rid of this effect, is to divide out each gene by its average expression within the set. effectivly amplifying the differences between libraries. I will do this first, and then get into more nuanced normalization schemes. 

```{r scatter_basic_norm1, fig.width=11, fig.height=11, cache=TRUE}
check_stress_mean_normalized = function(stress) {
  stress_meta = master_meta %>% filter(Strain=='WT', Stress==stress)
  set = log2(p[,stress_meta$Sample_Name])
  set_mean = log2(rowMeans(p[,stress_meta$Sample_Name]))
  set_norm = apply(set, 2, function(x) x - set_mean) #Remeber that subtraction in log space is the same as division before, so here I am dividing each probablility by the mean probablility
  pairs(set_norm,pch=20, xlim=c(-3.5,3.5), ylim=c(-3.5,3.5))
}
check_stress_mean_normalized('Menadione')
check_stress_mean_normalized('Azetidine')
check_stress_mean_normalized('Heatshock')
```
This is a bad idea, becuase if the data is all identical, we would get scatter plots all on the zero,zero line. 

what we would rather have is the baseline be the average for ideal condtions. 

So lets try that:


```{r scatter_basic_norm2, fig.width=11, fig.height=11, cache=TRUE}
check_stress_mean_normalized = function(stress) {
  stress_meta = master_meta %>% filter(Strain=='WT', Stress==stress)
  control_meta = master_meta %>% filter(Strain=='WT', Stress=='None')
  set = log2(p[,stress_meta$Sample_Name])
  set_mean = log2(rowMeans(p[,control_meta$Sample_Name]))
  set_norm = apply(set, 2, function(x) x - set_mean) #Remeber that subtraction in log space is the same as division before, so here I am dividing each probablility by the mean probablility
  pairs(set_norm,pch=20)#, xlim=c(-3.5,3.5), ylim=c(-3.5,3.5))
}
check_stress_mean_normalized('Menadione')
check_stress_mean_normalized('Azetidine')
check_stress_mean_normalized('Heatshock')
```

Now I will attempt to remove the base-line expression level effect from this data. But remember how well the correlations looked within group. now removing the baseline effects, whatever score they should produce, should be cosidered a good score. 

I will do the normalization on a group by group basis. 

Each of these groups have some sort of baseline expression. for some of these experiments there are further Caviets:
1. For Xu's menadion data:
..* His 3 repeats of WT in YPD. which I will average (but I also could look at each individually)
..* These were done on the same day as his menadione samples
2. For my menadione set with drug, m:
..* 4 repeats with drug and in tubes,
..* The baseline in this set were done on differnt days, it is possible that the cells were at a differnt phase
3. For my menadione set _without_ drug (the K set):
..* 4 WT strains just in plates and in YPD. 
..* This baseline was done on a different day

```{r scatter_basic_norm, fig.width=11, fig.height=11}
# This is a little hacky, because im nut sure how to get it done elegantly
men = data.frame(
  Xu_1 = p[,"WT_MEN1"]/p[,"WT_C1"],
  Xu_2 = p[,"WT_MEN2"]/p[,"WT_C2"],
  Xu_3 = p[,"WT_MEN3"]/p[,"WT_C3"],
  Drug_25 = p[,"25m"]/p[,"25yd"],
  Drug_26 = p[,"26m"]/p[,"26yd"],
  Drug_27 = p[,"27m"]/p[,"27yd"],
  Drug_28 = p[,"28m"]/p[,"28yd"],
  No_Drug_25 = p[,"4k"]/p[,"25yc"],
  No_Drug_26 = p[,"12k"]/p[,"26yc"],
  No_Drug_27 = p[,"20k"]/p[,"27yc"],
  No_Drug_28 = p[,"28k"]/p[,"28yc"])
pairs(log2(men),xlim=c(-8,8),ylim=c(-8,8))

azc = data.frame(
  Xu_1 = p[,"WT_AZC1"]/p[,"WT_C1"],
  Xu_2 = p[,"WT_AZC2"]/p[,"WT_C2"],
  Xu_3 = p[,"WT_AZC3"]/p[,"WT_C3"],
  Drug_25 = p[,"25a"]/p[,"25yd"],
  Drug_26 = p[,"26a"]/p[,"26yd"],
  Drug_27 = p[,"27a"]/p[,"27yd"],
  Drug_28 = p[,"28a"]/p[,"28yd"],
  No_Drug_25 = p[,"6k"]/p[,"25yc"],
  No_Drug_26 = p[,"14k"]/p[,"26yc"],
  No_Drug_27 = p[,"22k"]/p[,"27yc"],
  No_Drug_28 = p[,"30k"]/p[,"28yc"])
pairs(log2(azc),xlim=c(-8,8),ylim=c(-8,8))

# Did Joana do 1NM with her WT_HS ? its important for the control, either use WT_C or WT_1NM
heat = data.frame(
  Xu_1 = p[,"WT_HS1"]/p[,"WT_C1"],
  Xu_2 = p[,"WT_HS2"]/p[,"WT_C2"],
  Xu_3 = p[,"WT_HS3"]/p[,"WT_C3"],
  Joana = p[,"WT_HS"]/p[,"WT_C"],
  Drug_25 = p[,"25h"]/p[,"25yd"],
  Drug_26 = p[,"26h"]/p[,"26yd"],
  Drug_27 = p[,"27h"]/p[,"27yd"],
  Drug_28 = p[,"28h"]/p[,"28yd"],
  No_Drug_25 = p[,"7k"]/p[,"25yc"],
  No_Drug_26 = p[,"15k"]/p[,"26yc"],
  No_Drug_27 = p[,"23k"]/p[,"27yc"],
  No_Drug_28 = p[,"31k"]/p[,"28yc"])
pairs(log2(heat),xlim=c(-8,8),ylim=c(-8,8))
```



just for fun:

```{r}
ggobi(log2(heat))
ggobi(log2(azc))
ggobi(log2(men))
```

## Conclusion
In conclusion, the effect caused by the stress is still evident in samples, however the plate format, coupled with the drug also contribute signal that is on the same order of magnitude as the stress. this makes the 'condition' more than just the stress. this makes comparing these conditions with stess conditions from other sources problematic.

## Future Directions
1. Do this same analysis, but use DESeq to normalize the log counts such that the lowly counted genes dont mess with the data
2. Look at DE genes in the given condition only. what if we restrict our data to only looking at those genes?
