---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---

One question we want to know is: what genes are differentially expressed between our kinase strains and the WT strains regardless of condition. for a given kinase, there will be 3 classes of genes. 

1. Those whos levels do not differ from WT ever
2. Those whos levels differ from WT only in certain conditions
3. Those whos levels differ from WT in all conditions

Differentating from class 1 and class 3 seems to be doable. class 2 genes may be classified as class 1 genes if the number of conditions are small (statistcally seen as an outlier), but will be classified as class 3 if the number of conditions that have differential expression with respect to WT.

```{r, message=FALSE}
source('../utils/load_libraries.R')
source('../utils/load_data.R')
source('../utils/load_functions.R')
library(ggrepel)
```

```{r}


drug_counts = raw_counts[,meta$Drug == 'Cocktail']
drug_meta = meta[meta$Drug == 'Cocktail', ]
dds = DESeqDataSetFromMatrix(countData = drug_counts, 
                             colData = drug_meta, 
                             design = ~ Stress + Strain)
dds = DESeq(dds, parallel = TRUE)
#colnames(dds) = 
#dds = dds[,dds$Drug == 'Cocktail']
#dds = DESeq(dds, test="LRT", reduced = ~ Stress, parallel = TRUE)

#res = results(dds)
#res
#summary(res)


```
Lets do a wald test for hog1
```{r}
volcano = function(dds, contrast, t2g, write_genes = TRUE) {
  res = results(dds, contrast=c("Strain",contrast,"WT"))
  res$Gene = t2g[match(rownames(res), t2g$target_id),]$name
  res$Description = t2g[match(rownames(res), t2g$target_id),]$description

dat = as.data.frame(res) %>% 
  filter(!is.na(padj) & Gene != 'HIS3' & Gene != 'YOR203W') %>% 
  mutate(Significant = padj < 0.01, 
         Large_Difference = abs(log2FoldChange) > 1, 
         Sig_Diff = factor(paste(Significant, Large_Difference, sep="_"),
                           levels = c('FALSE_FALSE', 'FALSE_TRUE', 'TRUE_FALSE', 'TRUE_TRUE'),
                           labels = c('None', 'Large Difference', 'Significant', 'Large and Significant')))
dat$Gene[dat$Sig_Diff != 'Large and Significant'] = NA

if(write_genes){
  ouput_dir = '../../output/Strain_Dependent_Genes/'
  write(dat$Gene[dat$Sig_Diff == 'Large and Significant'], 
        file = paste0(ouput_dir, contrast, '_sig.txt'))
  write(dat$Gene[dat$Sig_Diff == 'Large and Significant' & dat$log2FoldChange > 0],
        file = paste0(ouput_dir, contrast, '_sig_up.txt'))
  write(dat$Gene[dat$Sig_Diff == 'Large and Significant' & dat$log2FoldChange < 0],
        file = paste0(ouput_dir, contrast, '_sig_down.txt'))
}


if(table(dat$Sig_Diff)[4] > 200){
  dat$Gene = NA
}

plot = ggplot(dat,
              aes(
                x = log2FoldChange,
                y = -log10(padj),
                color = Sig_Diff,
                label = Gene
              )) + geom_point() +
  geom_text_repel(na.rm = TRUE, size=3) + 
  ggtitle(contrast)
ggsave(filename=paste0('../../output/Strain_Dependent_Genes/', contrast, '.pdf'),
       plot=plot)

return(plot)
}

#volcano(dds, 'HOG1', t2g)
```

```{r}
t2g = load_transcripts_to_genes()

```

```{r}
# library(calibrate)
# with(res, plot(log2FoldChange, -log10(pvalue), pch=20, main="Volcano plot", xlim=c(-1,2.5), ylim = c(0,40)))
# 
# # Add colored points: red if padj<0.05, orange of log2FC>1, green if both)
# with(subset(res, padj<.05 ), points(log2FoldChange, -log10(pvalue), pch=20, col="red"))
# with(subset(res, abs(log2FoldChange)>1), points(log2FoldChange, -log10(pvalue), pch=20, col="orange"))
# with(subset(res, padj<.05 & abs(log2FoldChange)>1), points(log2FoldChange, -log10(pvalue), pch=20, col="green"))
# 
# # Label points with the textxy function from the calibrate plot
# library(calibrate)
# with(subset(res, padj<.05 & abs(log2FoldChange)>1), textxy(log2FoldChange, -log10(pvalue), labs=Gene, cex=.8))

dat = as.data.frame(res) %>% 
  filter(!is.na(padj) & Gene != 'HIS3') %>% 
  mutate(Significant = padj < 0.05, 
         Large_Difference = abs(log2FoldChange) > 1, 
         Sig_Diff = factor(paste(Significant, Large_Difference, sep="_"), 
                           labels = c('None', 'Large Difference', 'Significant', 'Large and Significant')))

```


```{r}
# this is for the gene HUG1
#barchart(r['YML058W-A',])
```

```{r}
library(ggvis)
dat$id <- 1:nrow(dat)  # Add an id column to use ask the key

all_values <- function(x) {
  if(is.null(x)) return(NULL)
  row <- dat[dat$id == x$id, ]
  paste0("<b>", names(row), "</b>:", format(row), collapse = "<br />")
}

dat %>% 
  ggvis(x = ~log2FoldChange, y = ~-log10(padj), fill = ~Sig_Diff, text := ~Gene, key := ~id) %>%
  #layer_points() %>%
  layer_text(fontSize := 4) %>%
  add_tooltip(all_values, "hover")

```


Okay lets loop over parametrs and write to dir

```{r}
write_sig = function(dds, contrast, t2g, method = 'cutoff', cutoff=1, top = 100, pval = 0.01) {
  res = results(dds, contrast=c("Strain",contrast,"WT"))
  res$Gene = t2g[match(rownames(res), t2g$target_id),]$name
  res$Description = t2g[match(rownames(res), t2g$target_id),]$description

if(method == 'top'){
  fc = res$log2FoldChange[res$padj<pval]
  fc = fc[!is.na(fc)]
  cutoff_low = min(0, sort(fc)[top + 1])
  cutoff_high = max(0, sort(fc)[length(fc) - top])
} else {
  cutoff_low = cutoff
  cutoff_high = cutoff
}
  
dat = as.data.frame(res) %>% 
  filter(!is.na(padj) & Gene != 'HIS3' & Gene != 'YOR203W') %>% 
  mutate(Significant = padj < pval, 
         Large_Difference_Up = log2FoldChange > cutoff_high,
         Large_Difference_Down = log2FoldChange < cutoff_low)

  ouput_dir = '../../output/Strain_Dependent_Genes/'
  file_root = paste0(ouput_dir, 
                     contrast,'_', 
                     pval, '_', 
                     method, '_', 
                     if_else(method == 'top', top, cutoff))
  
  write(dat$Gene[dat$Significant & dat$Large_Difference_Up],
        file = paste0(file_root, '_sig_up.txt'))
  write(dat$Gene[dat$Significant & dat$Large_Difference_Down],
        file = paste0(file_root, '_sig_down.txt'))
}

kinase = as.character(unique(as.character(drug_meta$Strain)))
kinase = kinase[-7]

tops = c(20, 50, 100, 200)
cuts = c(0.3, 0.5, 0.8, 1)
pvals = c(1, 0.1, 0.05, 0.01)

tops_settings = expand.grid(top = tops, pval = pvals, contrast = kinase)
cuts_settings = expand.grid(cutoff = cuts, pval = pvals, contrast = kinase)

tops_settings$contrast = as.character(tops_settings$contrast)
cuts_settings$contrast = as.character(cuts_settings$contrast)

# tops
mapply ("write_sig", 
        top = tops_settings$top, 
        pval = tops_settings$pval, 
        contrast = tops_settings$contrast,
        MoreArgs=list(dds = dds, t2g = t2g, method = 'top') )

# cuts
mapply ("write_sig", 
        cutoff = cuts_settings$cutoff, 
        pval = cuts_settings$pval, 
        contrast = cuts_settings$contrast,
        MoreArgs=list(dds = dds, t2g = t2g, method = 'cutoff') )


```