---
title: "Two Directions of Major Gene Expression"
output:
  html_document: default
  html_notebook: default
---

```{r,message=FALSE, echo=FALSE}
source('../utils/load_libraries.R')
```

```{r, message=FALSE, echo=FALSE, cache=TRUE}
source('../utils/load_functions.R')
source('../utils/load_data.R')
```

Let's look at the samples that have been inhibited:

```{r}
# Filter
vsd = vsd[,colData(vsd)$Drug == 'Cocktail']
meta = colData(vsd)
meta = as.data.frame(meta)
meta = meta %>% mutate(Condition = if_else(Stress=='None',                                                             as.character(Media),
                                           as.character(Stress)
                                          ))
meta$Condition = as.factor(meta$Condition)
x = assay(vsd)
colnames(x) = meta$Sample_Name
```

Let us also filter, keeping most, but increaseing the variance:

```{r}
rv = genefilter::rowVars(x)
select = order(rv, decreasing = TRUE)[seq_len(4000)]
x = x[select, ]
```

Let's now look at how the Principal Components of the genes separate the samples:

```{r}
pca = prcomp(t(x))
meta_pc = cbind(as.data.frame(meta),
                as.data.frame(
                  pca$x[match(meta$Sample_Name,rownames(pca$x)),]
                  ))

p1 = ggplot(meta_pc, aes(x=PC1, y=PC2))
p1 + geom_point(shape = 21, aes(fill = Condition), size=4, stroke=1) + geom_text(aes(label=Strain), size=1)
```

We see that the first two principal components create a triangular scatter, with 3 conditions in the corners:

* No Stress
* Glucose Dropout
* Rapamycin

Lets look at these 3 conditions:

```{r}
  for(s in c('YPD', 'Glucose Depletion', 'Rapamycin')) {
  plot = p1 + 
    geom_point(data = subset(meta_pc, Condition == s), 
                  aes(fill = Condition),
                  shape = 21, 
                  size=4, stroke=1) + 
    geom_text(aes(label=Strain), 
              size=1) + 
    scale_fill_discrete(drop = FALSE) +
    labs(title = s)  
  print(plot)
  }
  
```

This is interesting, Not only are these conditions the most extream in terms of distance from No Stress, they are also considered the cononical activators of two major Kinases: PKA and TOR.

I therefore hypothesize that these directions in PC space represent the PKA, and TOR pathways.

If this is true, I might expect that for each individual condition, the PKA and TOR2 strains would be offset from the group in the same direction as these pathways. lets check that out:

```{r}
  for(s in levels(meta$Condition)) {
  plot = p1 + 
    geom_point(data = subset(meta_pc, Condition == s), 
                  aes(fill = Condition,
                      size = if_else(Strain %in% c('TPK123','TOR2'),4,2),
                      color = if_else(Strain %in% c('TPK123','TOR2'),2,1)),
                  shape = 21, 
                  stroke=1) + 
    geom_text(aes(label=Strain), 
              size=1) + 
    scale_fill_discrete(drop = FALSE) +
    labs(title = s) +
    guides(size=FALSE, color=FALSE) +
    scale_size(range = c(3, 7))
  print(plot)
  }
  
```
  
  
One thing we definitely need to look at is how much of the variance is explained in the first two PCs? are we talking the whole thing? or just a part of it?

```{r}
plot(summary(pca)$importance[3,], 
     main='Percentage of Variance Explained',
     xlab='PCs',
     ylab='Cumulative Percentage of Variance Explained',
     type='l')

plot(summary(pca)$importance[3,1:5], 
     main='Percentage of Variance Explained',
     xlab='PCs',
     ylab='Cumulative Percentage of Variance Explained',
     type='l',
     ylim=c(0,1))
```
