---
title: "Global Analysis of Kinase Drivers of the Yeast Stress Response"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

```{r,message=FALSE, echo=FALSE}
source('../utils/load_libraries.R')
```

```{r, message=FALSE, echo=FALSE, cache=FALSE}
source('../utils/load_functions.R')
source('../utils/load_data.R')
```
# Setup
We will be condsidering the samples that have been inhibited:
```{r}
# Filter
vsd_no_drug = vsd[,colData(vsd)$Drug == 'None']
vsd = vsd[,colData(vsd)$Drug == 'Cocktail']
meta = colData(vsd)
meta = as.data.frame(meta)
meta = meta %>% mutate(Condition = if_else(Stress=='None',
                                           as.character(Media),
                                           as.character(Stress)
                                          ))
meta$Condition = as.factor(meta$Condition)
x = assay(vsd)
x = x - rowMeans(x)
colnames(x) = meta$Sample_Name

gasch = load_gasch()
gx = gasch$data
gmeta = gasch$meta

pronk = load_pronk()
px = pronk$data
pmeta = pronk$meta

mega = load_megaYeast()
mx = mega$data

common_genes = Reduce(intersect, 
                      list(rownames(gx), 
                           rownames(x), 
                           rownames(px),
                           rownames(mx)))

x = x[common_genes, ]
gx = gx[common_genes, ]
px = px[common_genes, ]
mx = mx[common_genes, ]
```
# Overview
##Prinicipal Component Decomposition
```{r}
pca = prcomp(t(x))
meta_pc = cbind(as.data.frame(meta),
                as.data.frame(
                  pca$x[match(meta$Sample_Name,rownames(pca$x)),]
                  ))
```
First, let look at how much variance is explained by each PC:
```{r}
barplot(summary(pca)$importance[2,1:10],
     main='Percentage of Variance Explained',
     xlab='PCs',
     ylab='Percentage of Variance Explained',
     ylim=c(0,1))
100 * summary(pca)$importance[2,1:9]
```
PC1 with 35%, PC2 with 20%, and the next PCs - 3 & 4 are down to ~7% of the variance.

Let's now look at how the first two Principal Component separate the samples:
```{r}
plot = ggplot(meta_pc, aes(x=PC1, y=PC2)) +
       geom_point(shape = 21, aes(fill = Condition, color=Drug), size=4, stroke=1) +
       geom_text(aes(label=Strain), size=1)

ggplotly(plot)
```
## Sample Extreams
The above plot has a intersting shape, a triangle with 3 conditions strongly defining the corners:
```{r}
plot = ggplot(meta_pc %>% filter(Condition %in% c('YPD','Glucose Depletion', 'Rapamycin')), aes(x=PC3, y=PC2)) +
       geom_point(shape = 21, aes(fill = Condition), size=4, stroke=1) +
       geom_text(aes(label=Strain), size=1) +
       scale_fill_discrete(drop = FALSE)

ggplotly(plot)
```
The conditions that make up the vertices of the triangle are:
* No Stress
* Glucose Dropout
* Rapamycin

These 3 are particularly intersting:
* No Stress could be considered an origin
* Rapamycin is the canonical activator of the TOR pathway
* Glucose Depletion is the cononical activators of the PKA pathway

#Hypothesis
## Statement
One potential hypothesis is therefore:

1. The direction of the two vertices away from the control indicate the pathway activity for PKA and TOR.
  + I will refer to these directions as the PKA and TOR __cardinal__ directions
2. As far as PKA and TOR go, every stress I tested is some linear combination of PKA and TOR, which could be used to create a hierachy of stresses

## Possible Evidence
If this hypothesis is true, I would expect the following:
1. Within each individual condition, activation of PKA and TOR2 pathways should move the sample in their respective __cardinal__ direction. 
2. If I project other stress datasets onto these prinicpal components, I should see a similar pattern.
3. If I look at the gene loadings of these cardinal directions, I should extract genes associated with PKA and TOR.

# Results
## Cardinality
```{r}
  for(s in levels(meta$Condition)) {
  plot = ggplot(meta_pc, aes(x=PC1, y=PC2)) +
         geom_point(data = subset(meta_pc, Condition == s),
                    aes(fill = Condition,
                    size = if_else(Strain %in% c('TPK123','TOR2'),4,2),
                    color = if_else(Strain %in% c('TPK123','TOR2'),2,1)),
                    shape = 21,
                    stroke=1) +
         geom_text(aes(label=Strain),
                   size=1) +
         scale_fill_discrete(drop = FALSE) +
         labs(title = s) +
         guides(size=FALSE, color=FALSE) +
         scale_size(range = c(3, 7))
  print(plot)
  }

```
<!-- If you want to use facet_wrap 
```{r, fig.height=14, fig.width=7}

  plot = ggplot(meta_pc, aes(x=PC1, y=PC2)) +
         geom_point(aes(fill = Condition,
                    size = if_else(Strain %in% c('TPK123','TOR2'),4,2),
                    color = if_else(Strain %in% c('TPK123','TOR2'),2,1)),
                    shape = 21,
                    stroke=1) +
         geom_text(aes(label=Strain),
                   size=1) +
         scale_fill_discrete(drop = FALSE) +
         guides(size=FALSE, color=FALSE) +
         scale_size(range = c(3, 7)) +
         facet_wrap(~Condition, scales = 'fixed', ncol = 2)
  print(plot)


```
-->

## Projection of other data
### Gasch
```{r}
g_on_x = predict(pca,t(gx))
gmeta_on_x = cbind(gmeta, g_on_x)

plot = ggplot(gmeta_on_x, 
              aes(x=PC3,
                  y=PC2,
                  label = ifelse(is.na(Time), "", sprintf("%1.0f",Time)),
                  fill = Stress_after)) + 
       geom_point(shape = 21, size=4, stroke=1) +
       geom_text(na.rm = FALSE, size=1.5)
ggplotly(plot, tooltip = c("y", "x", "Time", 'fill'))
# # This is to look at the PCs of the actual datset:
# gpca = prcomp(t(gx))
# gmeta_pc = cbind(as.data.frame(gmeta),
#                 as.data.frame(
#                   gpca$x[match(gmeta$Sample_Name,rownames(gpca$x)),]
#                   ))
# plot = ggplot(gmeta_pc, aes(x=PC1, y=PC2)) + geom_point(shape = 21, size=4, stroke=1)
# print(plot)

# # To include time:
# plot = ggplot(gmeta_pc, aes(x=PC1, y=PC2)) +
#        geom_point(shape = 21, 
#                   aes(fill = Stress_after), 
#                   size=4, 
#                   stroke=1) + 
#        geom_text(aes(label = ifelse(is.na(Time), "", sprintf("%1.0f",Time))), 
#                  na.rm = FALSE, 
#                  size=2)
```
### Pronk

```{r}
p_on_x = predict(pca,t(px))
pmeta_on_x = cbind(pmeta, p_on_x)

plot = ggplot(pmeta_on_x, 
              aes(x=PC1,
                  y=PC2,
                  fill = Condition)) + # TODO Add color when you have meta 
       geom_point(shape = 21, size=4, stroke=1) #+
       #geom_text(na.rm = FALSE, size=1.5)
ggplotly(plot)#, tooltip = c("y", "x", "Time", 'fill'))

# # This is to look at the PCs of the actual datset:
# ppca = prcomp(t(px))
# pmeta_pc = cbind(as.data.frame(pmeta),
#                 as.data.frame(
#                   ppca$x[match(pmeta$title,rownames(ppca$x)),]
#                   ))
# plot = ggplot(pmeta_pc, aes(x=PC1, y=PC2)) + geom_point(shape = 21, size=4, stroke=1)
# print(plot)
```



```{r}
m_on_x = predict(pca,t(mx))
#pmeta_on_x = cbind(pmeta, p_on_x)

plot = ggplot(data.frame(m_on_x), 
              aes(x=PC1,
                  y=PC2)) + # TODO Add color when you have meta 
       geom_point(shape = 21, size=4, stroke=1) #+
       #geom_text(na.rm = FALSE, size=1.5)
ggplotly(plot)#, tooltip = c("y", "x", "Time", 'fill'))

# # This is to look at the PCs of the actual datset:
# ppca = prcomp(t(px))
# pmeta_pc = cbind(as.data.frame(pmeta),
#                 as.data.frame(
#                   ppca$x[match(pmeta$title,rownames(ppca$x)),]
#                   ))
# plot = ggplot(pmeta_pc, aes(x=PC1, y=PC2)) + geom_point(shape = 21, size=4, stroke=1)
# print(plot)
```






Can we compare the PCs of each dataset?

```{r}
pca = prcomp(t(x))
gpca = prcomp(t(gx))
ppca = prcomp(t(px))
mpca = prcomp(t(mx))

gsim = cor(pca$rotation, gpca$rotation)
psim = cor(pca$rotation, ppca$rotation)
msim = cor(pca$rotation, mpca$rotation)

get_best_pc = function(source_pc, target_pc, num_target, num_best_source){
  sim = cor(source_pc$rotation, target_pc$rotation)
  abs_sim = abs(sim)
  care = abs_sim[,1:num_target]
  source_pc_names = rownames(care)
  best = apply(care,2,function(x) source_pc_names[order(x,decreasing = T)[1:num_best_source]])
  return(best)
}
#get_best_pc(pca,pca,10,5)
```
What are the best PCs from my data to match Gasch PCs?
```{r}
get_best_pc(pca,gpca,10,3)
```
What are the best PCs from my data to match Pronk PCs?
```{r}
get_best_pc(pca,ppca,10,3)
```
What are the best PCs from my data to match Mega PCs?
```{r}
get_best_pc(pca,mpca,10,3)
```

Two major questions I want to answer are:

What are each PC indiciating? lets make a way to extract go terms for each PC.

For a given stress, which strain is an outlier in PC space

What is the cardinality for each strain across all stresses? does inactivaton of a particualr stress always point in the same direction in PC space? and if so, what direction does it point in? are there go terms associated with that direction?

PBS2 seems really intersting. it seems to shut down the glucose dropout response. which itself is a very strong response. what genes is it shutting down? the entire program? or part of it (such as msn2 targets or tod6 targets etc. )

Lets describe the PCs using go terms.

```{r, results='asis', message=F}
get_go_terms_from_list = function(query_list, background_list) {
    library(goseq)
    library(org.Sc.sgd.db)
    genes = numeric(length(background_list))
    names(genes) = background_list
    genes[query_list] = 1
    pwf = nullp(genes,"sacCer2","ensGene")
    GO.wall=goseq(pwf,gene2cat=as.list(org.Sc.sgdGO2ALLORFS))
    return(GO.wall)
}

print_out_terms = function(g_dir){
  library(knitr)
  g_dir %>%
    dplyr::mutate(pval = min(over_represented_pvalue, under_represented_pvalue),
                  percent_in = numDEInCat / numInCat) %>%
    dplyr::arrange(pval) %>%
    dplyr::filter(pval<0.01) %>% 
    dplyr::select(term, pval, percent_in) %>% head() %>% kable(format = "markdown")
}

do_pc = function(str, pc){
pc = pca$rotation[,str]

pc_up = names(pc)[(pc - mean(pc))/sd(pc) > 2]
g_up = get_go_terms_from_list(pc_up, names(pc))
print_out_terms(g_up)

pc_down = names(pc)[(pc - mean(pc))/sd(pc)< -2]
g_down = get_go_terms_from_list(pc_down, names(pc))
print_out_terms(g_down)
}

do_pc('PC1', pca)
do_pc('PC2', pca)
do_pc('PC3', pca)
do_pc('PC4', pca)

```

