---
title: "Hypothesis: General Stress Response is primarily a linear combination of two programs"
output:
  html_document: default
  html_notebook: default
---

```{r,message=FALSE, echo=FALSE}
source('../utils/load_libraries.R')
```

```{r, message=FALSE, echo=FALSE, cache=TRUE}
source('../utils/load_functions.R')
source('../utils/load_data.R')
```

# General Overview

Let's look at the samples that have been inhibited:

```{r}
# Filter
vsd = vsd[,colData(vsd)$Drug == 'Cocktail']
meta = colData(vsd)
meta = as.data.frame(meta)
meta = meta %>% mutate(Condition = if_else(Stress=='None',                                                             as.character(Media),
                                           as.character(Stress)
                                          ))
meta$Condition = as.factor(meta$Condition)
x = assay(vsd)
colnames(x) = meta$Sample_Name
```

Let us also filter, keeping most, but increaseing the variance:

```{r}
rv = genefilter::rowVars(x)
select = order(rv, decreasing = TRUE)[seq_len(4000)]
x = x[select, ]
```

# Prinicipal Component Decomposition

Let's now look at how the Principal Components of the genes separate the samples:

```{r}
pca = prcomp(t(x))
meta_pc = cbind(as.data.frame(meta),
                as.data.frame(
                  pca$x[match(meta$Sample_Name,rownames(pca$x)),]
                  ))

p1 = ggplot(meta_pc, aes(x=PC1, y=PC2))
p1 + geom_point(shape = 21, aes(fill = Condition), size=4, stroke=1) + geom_text(aes(label=Strain), size=1)
```

We see that the first two principal components create a triangular scatter, with 3 conditions in the corners:

* No Stress
* Glucose Dropout
* Rapamycin

Lets look at these 3 conditions:

```{r}
  for(s in c('YPD', 'Glucose Depletion', 'Rapamycin')) {
  plot = p1 + 
    geom_point(data = subset(meta_pc, Condition == s), 
                  aes(fill = Condition),
                  shape = 21, 
                  size=4, stroke=1) + 
    geom_text(aes(label=Strain), 
              size=1) + 
    scale_fill_discrete(drop = FALSE) +
    labs(title = s)  
  print(plot)
  }
  
```

This is interesting, Not only are these conditions the most extream in terms of distance from No Stress, they are also considered the cononical activators of two major Kinases: PKA and TOR.

I therefore hypothesize that these directions in PC space represent the PKA, and TOR pathways.

If this is true, I might expect that for each individual condition, the PKA and TOR2 strains would be offset from the group in the same direction as these pathways. lets check that out:

```{r}
  for(s in levels(meta$Condition)) {
  plot = p1 + 
    geom_point(data = subset(meta_pc, Condition == s), 
                  aes(fill = Condition,
                      size = if_else(Strain %in% c('TPK123','TOR2'),4,2),
                      color = if_else(Strain %in% c('TPK123','TOR2'),2,1)),
                  shape = 21, 
                  stroke=1) + 
    geom_text(aes(label=Strain), 
              size=1) + 
    scale_fill_discrete(drop = FALSE) +
    labs(title = s) +
    guides(size=FALSE, color=FALSE) +
    scale_size(range = c(3, 7))
  print(plot)
  }
  
```
  
  
One thing we definitely need to look at is how much of the variance is explained in the first two PCs? are we talking the whole thing? or just a part of it?

```{r}
plot(summary(pca)$importance[3,], 
     main='Percentage of Variance Explained',
     xlab='PCs',
     ylab='Cumulative Percentage of Variance Explained',
     type='l')

plot(summary(pca)$importance[3,1:5], 
     main='Percentage of Variance Explained',
     xlab='PCs',
     ylab='Cumulative Percentage of Variance Explained',
     type='l',
     ylim=c(0,1))

barplot(summary(pca)$importance[2,1:10], 
     main='Percentage of Variance Explained',
     xlab='PCs',
     ylab='Percentage of Variance Explained',
     ylim=c(0,1))
```

So we can see that the first two give big gains. 

```{r}
100 * summary(pca)$importance[2,1:9]
```

PC1 with 35%, PC2 with 20%, and the next PCs - 3 & 4 are down to ~7%


# Looking at Gasch Data

```{r}
gasch = read.table('../../input/reference/gasch/pickGenes/imputed_gasch_data.txt', header=T, row.names = 1)
gmeta = read.csv('../../input/meta/gasch_meta.csv')
Stress_after = as.character(gmeta$Stress)
Stress_after[gmeta$Time<10] = 'Pre-response'
gmeta$Stress_after = factor(Stress_after)
x = gasch
rv = genefilter::rowVars(x)
select = order(rv, decreasing = TRUE)[seq_len(4000)]
x = x[select, ]
colnames(x) = gmeta$Sample_Name
pca = prcomp(t(x))
meta_pc = cbind(as.data.frame(gmeta),
                as.data.frame(
                  pca$x[match(gmeta$Sample_Name,rownames(pca$x)),]
                  ))

p1 = ggplot(meta_pc, aes(x=PC1, y=PC2))
p1 + geom_point(shape = 21, aes(fill = Stress_after), size=4, stroke=1) + geom_text(aes(label = ifelse(is.na(Time), "", sprintf("%1.0f",Time))), na.rm = FALSE, size=2) 

```



```{r}
all_values <- function(x) {
  if(is.null(x)) return(NULL)
  row <- meta_pc[meta_pc$Sample_Name == x$Sample_Name, ]
  paste0(names(row[1:2]), ": ", format(row[1:2]), collapse = "<br />")
  #return(x$Sample_Name)
}

meta_pc %>% 
   ggvis(x = ~PC1,
         y = ~PC3,
         key := ~Sample_Name) %>% 
   layer_points(fill = ~factor(Stress)) %>%
   add_tooltip(all_values, "hover")
   #add_tooltip(function(x) paste0(as.character(x$Stress), ' : ', x$Sample_Name, collapse = "<br />"), "hover")
   
```

Now lets also look at the PCs

```{r}
plot(summary(pca)$importance[3,], 
     main='Percentage of Variance Explained',
     xlab='PCs',
     ylab='Cumulative Percentage of Variance Explained',
     type='l')

plot(summary(pca)$importance[3,1:5], 
     main='Percentage of Variance Explained',
     xlab='PCs',
     ylab='Cumulative Percentage of Variance Explained',
     type='l',
     ylim=c(0,1))

barplot(summary(pca)$importance[2,1:10], 
     main='Percentage of Variance Explained',
     xlab='PCs',
     ylab='Percentage of Variance Explained',
     ylim=c(0,1))

100 * summary(pca)$importance[2,1:9]



```