Load up data and colors

```{r}
library(tidyverse)
library(ComplexHeatmap)
library(circlize)
load('../../../intermediate/images/paper_data.RData')
load('../../../intermediate/images/externally_defined_modules.RData')
source('make_obj.R')
source('colors.R')
library(ggrepel)
library(ggthemes)
```

# General Overview of data

```{r}

throw=read_csv('../../../intermediate/Samples_to_thow_out.csv') %>% left_join(meta)

normalized_counts %>% as_tibble(rownames='name') %>% gather(key = Sample_Name, value = norm_counts, -name) %>% mutate(log2Counts = log2(norm_counts + 1)) %>% group_by(name) %>% filter(mean(log2Counts)>5) %>% ungroup() %>% left_join(meta %>% select(Sample_Name, Strain, Strain_Code, Condition)) %>% mutate(Sample_Info = paste0(Strain_Code,'_',Condition)) %>% filter(Strain_Code != 'WT4' & !(Sample_Name %in% throw$Sample_Name)) %>% group_by(name) %>% mutate(deltaCount = log2Counts - mean(log2Counts[Strain == 'WT' & Condition == 'YPD'])) -> l2FC

l2FC %>% select(name,Sample_Info,deltaCount) %>% spread(key=Sample_Info,value = deltaCount) %>% remove_rownames() %>% column_to_rownames('name') %>% as.matrix() -> l2FC_matrix

l2FC %>% ungroup() %>% select(Sample_Info, Strain, Condition) %>% unique() %>% arrange(Condition, Strain, Sample_Info) -> l2FC_meta

l2FC_matrix = l2FC_matrix[,l2FC_meta$Sample_Info]

hm = Heatmap(l2FC_matrix[rownames(l2FC_matrix) %in% hs_sensitive,],
        col = colorRamp2(c(-4, -3, -2, -1, 0, 1, 2, 3, 4), coolwarm_hcl),
        show_row_names = F,
        show_column_names = F,
        cluster_columns = F,
        use_raster = T)
draw(hm)
```


# General Overview of computational methods
# Conditions and kinase relationships
# kinase, kinase relationships PBS2 PKA [general], HOG1 PBS2 [condition specific]
# classification of general and specific kinases
# kinase relationship with modules (rainbow network edges with colors for conditions)
# Fully connected relationships:
## residual 1 - Mating pathway
## residual 2 - proteasome with tor
# growth stuff - relationship between growth and expression

# Results:

```{r}
opts_knit$set(warning=FALSE, message=FALSE, error=FALSE)
```

```{r}

```
