---
title: "Data Processing"
output: html_notebook
---


This notebook takes us from the DESeq2 normalized data, to all the data that is used for plotting in the paper.

```{r data_loading}

rm(list=ls())

load('../../intermediate/images/normalized_data.RData')
load('../../intermediate/images/externally_defined_modules.RData')

library(tidyr)
library(purrr)
library(dplyr)
library(broom)
library(tibble)
library(magrittr)
library(forcats)
meta$Condition = fct_recode(meta$Condition, GlucoseDepletion = 'Glucose_Depletion')
raw_log = log2(raw_counts + 1)
measurements = raw_log %>% # vlog
  as.data.frame %>%
  rownames_to_column("Gene") %>%
  as_tibble %>%
  gather(Sample_Name, Expression, -Gene)

blind_measurements = vlog_blind %>%
  as.data.frame %>%
  rownames_to_column("Gene") %>%
  as_tibble %>%
  gather(Sample_Name, Blind_Expression, -Gene)

#rm(vlog)

meta = as_tibble(meta)

colnames(t2g)[1] = 'Gene'
t2g = as_tibble(t2g)

# should be a left join on the meta subsetted
meta_sub = meta %>% select(Sample_Name, Condition, Strain, Strain_Code)

measurements = measurements %>% left_join(blind_measurements) %>% left_join(meta_sub)

# Remove known bad sample

measurements %<>% filter(Strain_Code != 'WT1' | Condition != 'Rapamycin')

```

Next lets calculate some base statistics and define outliers:

```{r gene_statistics}

is_outlier <- function(v, coef=1.5){
  quantiles <- quantile(v,probs=c(0.25,0.75))
  IQR <- quantiles[2]-quantiles[1]
  res <- v < (quantiles[1]-coef*IQR)|v > (quantiles[2]+coef*IQR)
  return(res)
}


measurements = measurements %>% group_by(Gene, Condition) %>%
                mutate(is_condition_outlier = is_outlier(Expression),
                       condition_outlier = ifelse(is_condition_outlier,
                                                  as.character(Strain),
                                                  NA)) %>%
                ungroup() %>%
              group_by(Gene, Strain) %>%
                mutate(is_strain_outlier = is_outlier(Expression),
                       strain_outlier = ifelse(is_strain_outlier,
                                               as.character(Condition),
                                               NA)) %>%
                ungroup()

genes = measurements %>%
  group_by(Gene) %>%
  mutate(Differential_Expression = Expression - mean(Expression),
         Norm_Differential_Expression = Differential_Expression / sd(Expression)) %>%
  nest() %>%
  mutate(Gene_range = map_dbl(data, ~ diff(range(.$Expression))),
         Gene_sd = map_dbl(data, ~ sd(.$Expression)),
         Gene_mean = map_dbl(data, ~ mean(.$Expression))) %>%
  left_join(t2g)


```

create the linear model and extract model data:

```{r create_linear_models}

genes %<>%
  mutate(model = map(data, ~lm(Expression ~ Strain + Condition, data = .))) %>%
  mutate_at(.vars = vars(model),
            .funs = funs(performance = map(.,glance),
                         weights = map(.,tidy),
                         data_augment = map(.,augment)))

# models = genes %>%
#   left_join(modules) %>% select(data, module, name) %>% unnest() %>%
#   group_by(module) %>% nest() %>% sample_n(5) %>%
#   mutate(model = map(data, ~lm(Differential_Expression ~ Strain * Condition, data = .))) %>%
#   mutate_at(.vars = vars(model),
#             .funs = funs(performance = map(.,glance),
#                          weights = map(.,tidy),
#                          data_augment = map(.,augment)))


```

lets look at model performance, and figure out where our model breaks:

```{r model_performance}
library(ggplot2)
genes %>%
  select(Gene_mean, performance) %>%
  unnest() %>%
ggplot(aes(x=Gene_mean, y = r.squared)) +
  geom_point() +
  geom_vline(xintercept = 5) # TODO change this all to 5.5
```

It seems $R^2$ looks pretty bad when gene mean is below 5.5, probably becuase these genes are dominated by noise. lets therefore filter them out, and be save by filtering at 6:

```{r gene_filter}
genes %<>%
  filter(Gene_mean > 5) #TODO change this to 5.5 used to be mean > 6 to get all the good linear models. 3.6 seems good when looking at sd vs mean
```



Next lets do some basic clustering on the Differential Expression:

```{r tsne_cluster}
library(tsne)

do_tsne = function(mat){

  mat_meta = tibble(sample = colnames(mat)) %>%
    separate(sample, c('Condition', 'Strain_Code'), "_", remove = FALSE)

  pca = prcomp(t(mat))
  pca = data.frame(pca$x, sample = rownames(pca$x))
  meta_pc = full_join(mat_meta, pca)
  t = tsne(meta_pc %>% select(starts_with('PC')))
  colnames(t) = c('TSNE1', 'TSNE2')
  meta_pc = cbind(meta_pc, t)
  return(meta_pc)
}

gene_matrix = genes %>%
                unnest(data) %>%
                select(Gene, Condition, Strain_Code, Differential_Expression) %>%
                unite(sample, c(Condition, Strain_Code), remove = TRUE) %>%
                ungroup() %>%
                spread(sample, Differential_Expression)

sample_tsne = do_tsne(gene_matrix %>% select(-Gene))

```



Now lets extract the residuals for each gene model into it's own row in a wide matrix

```{r extract_residuals}
std_residuals = genes %>%
  select(data, data_augment, name) %>%
  unnest() %>%
  select(name, Condition, Strain_Code, .std.resid) %>%
  unite(sample_id, c(Strain_Code, Condition)) %>%
  rename(residual = .std.resid)

std_resid_matrix = std_residuals %>%
                     spread(key = sample_id,
                            value=residual,
                            drop = TRUE) %>%
                     remove_rownames() %>%
                     column_to_rownames(var = 'name') %>%
                     as.matrix()

raw_residuals = genes %>%
  select(data, data_augment, name) %>%
  unnest() %>%
  select(name, Condition, Strain_Code, .resid) %>%
  unite(sample_id, c(Strain_Code, Condition)) %>%
  rename(residual = .resid)

raw_resid_matrix = raw_residuals %>%
                     spread(key = sample_id,
                            value=residual,
                            drop = TRUE) %>%
                     remove_rownames() %>%
                     column_to_rownames(var = 'name') %>%
                     as.matrix()

```

Lets sort the genes into clusters by clustering on significantly large residuals: TODO decide if we should also include the weights as additional columns to cluster on?

Also, since small residuals are not significant, we will reduce any residual smaller than +/-2 to zero

```{r get_modules_external}

library(readr)
library(igraph)
library(gtools)
library(WGCNA)
library(dplyr)

yeastnet = read_delim('../../input/external_datasets/YeastNet.v3.txt', delim = '\t', col_names = c('Gene_1', 'Gene_2', 'weight'))
yeastnet %>% na.omit() %>% graph.data.frame(directed = FALSE) -> g
adj = as_adjacency_matrix(g, attr='weight', type='both')
adj = as.matrix(adj)
adj_inv_logit = inv.logit(adj)
adj_inv_logit = (adj_inv_logit*2 - 1)
diag(adj_inv_logit) = 1
p_est = pickSoftThreshold.fromSimilarity(similarity = adj_inv_logit)$powerEstimate
TOM = TOMsimilarity(adjMat = adj_inv_logit^p_est)
dissTOM = 1-TOM
# Call the hierarchical clustering function
geneTree = hclust(as.dist(dissTOM), method = "average");
# Plot the resulting clustering tree (dendrogram)


minModuleSize = 10;
# Module identification using dynamic tree cut:
dynamicMods = cutreeDynamic(dendro = geneTree, distM = dissTOM,
deepSplit = 2, pamRespectsDendro = FALSE,
minClusterSize = minModuleSize);

modules = data.frame(module = dynamicMods, 
                     Gene = rownames(adj_inv_logit)) %>%
  left_join(t2g)

# modules = tibble(module = as.character(net$colors),
#                  name = colnames(resid)) %>%
#                    group_by(module) %>%
#                    nest(.key = genes)

```




Lets get all the addtional information for each module
```{r eigengenes}
MEs = moduleEigengenes(t(vlog[modules$Gene[modules$Gene %in% rownames(vlog)],]), 
                 modules$module[modules$Gene %in% rownames(vlog)], 
                 impute = TRUE, 
                 nPC = 1, 
                 align = "along average", 
                 excludeGrey = FALSE, 
                 grey = if (is.numeric(colors)) 0 else "grey",
                 subHubs = TRUE,
                 softPower = 6,
                 scale = TRUE,
                 verbose = 0, indent = 0)$eigengenes

MEsCor = cor(MEs)


```

```{r module_avg_residual}
module_raw_residuals = raw_residuals %>% 
  left_join(modules) %>% 
  group_by(module, sample_id) %>% 
  summarize(residual = mean(residual)) %>% 
  ungroup()

module_std_residuals = std_residuals %>% 
  left_join(modules) %>% 
  group_by(module, sample_id) %>% 
  summarize(residual = mean(residual)) %>% 
  ungroup()
```

Lets get go enrichment for each module
```{r gsea}
# library("anRichment")
# 
# library(org.Sc.sgd)
# 
# GOenr = GOenrichmentAnalysis(dynamicMods, yeastORFs = colnames(resid), organism = 'yeast')
# module_terms = GOenr$bestPTerms[[4]]$enrichment


```

Now lets export the modules for meme analysis

```{r meme_prep}
cluster_dir = '../../intermediate/clusters/WGCNA'
output_dir = paste0(cluster_dir, '/gene_lists')
dir.create(output_dir, recursive = TRUE)

library(readr)

write_genes = function(df, output_dir){
  write(df$name,
        file = paste0(output_dir,
                      '/cluster_',
                      unique(df$module),
                      '.txt')
        )
}

modules %>%
  split(.$module) -> by_module

lapply(by_module, function(x) write_genes(x, output_dir))

```


  Run MEME

```{bash meme}
bash src/scripts/meme/run_all.sh intermediate/clusters/WGCNA/gene_lists intermediate/clusters/WGCNA

```





  # After meme is run:
```{r meme_import}

module_tfs = do.call(rbind,
lapply(dir(paste0(cluster_dir, '/tfs/')),
function(x) {
  path = paste0(cluster_dir, '/tfs/', x)
  if(file.size(path) > 0) {
    return(data.frame(file = x,
                      TF = read.table(path, stringsAsFactors = F))
           )
  }
})) %>%
 as_tibble %>%
 separate(file, c('c', 'module', 't'), sep = '_') %>%
 dplyr::rename(TF = V1) %>%
 dplyr::select(module, TF)
```



```{r data_saving}
save(list = c('std_resid_matrix',
              'std_residuals',
              'raw_resid_matrix',
              'raw_residuals',
              "module_raw_residuals",
              "module_std_residuals",
              "MEsCor",
              "MEs",
              'genes',
              'meta',
              'modules',
              'module_tfs',
              'sample_tsne',
              't2g'),
     file = '../../intermediate/images/paper_data.RData')
```

```{r relative_importance}
# library(relaimpo)
# getri = function(mod){
# r = calc.relimp(mod, type = c('last'))@last
# return(r)
# }
# lasts = do.call(rbind, lapply(models$Full_model, getri))
```
