## The full linear model of Condition and Stress identity accurately describes gene expression data
(or at least for most genes)

1. What's happening: I'm going to build the linear model
2. I'm going to show that it has good performance over most genes. 

### Building the model:

```{r, eval = FALSE}

```
So we've built a model for each gene. lets quickly look at the data frame 


```{r}
colnames(t2g)[1] = 'Gene'
genes = genes %>% left_join(t2g)
#knitr::kable(head(genes,100))
colnames(genes)
```

In order to asses the quality for each fit, lets calculate the $R^2$, or the percentage of variance explained for each gene from the model.

```{r}
mod_q = genes %>%
  select(Gene, name, data, ends_with('model')) %>%
  mutate_at(.vars=vars(ends_with('model')), 
            .funs = funs(glance = map(.,glance))) %>%
  select(Gene, name, ends_with('glance')) %>%
  gather(model_type, model_data, -Gene, -name) %>% 
  mutate(model_type = map(model_type, function(x) strsplit(x,'_')[[1]][1])) %>% 
  unnest()

head(mod_q)
```

Lets look at the $R^{2}$ for each gene in the full model:

```{r}
mod_q %>% 
    filter(model_type == 'Full') %>% 
    ggplot(aes(x=r.squared)) + 
    geom_density(fill='blue', alpha=.5) +
    ggtitle('Percentage of variance explained by the full model')
```
 
 Okay, so for most genes, the model does well. For what genes does it perform badly?
 
 
```{r}
mod_q %>% 
    filter(model_type == 'Full' & r.squared < .4) %>% 
    arrange(r.squared)
```
Lets look at an example of strong and week performace:

```{r}
library(stringr)

library(tidyr)
library(dplyr)
library(tibble)
library(ggplot2)
library(gridExtra)
library(RGraphics)
library(ggrepel)

expression_plot = funciton(data){
  
}

residual_plot = function(gene, all) {
  dat = all %>% dplyr::filter(name == gene) %>% unnest()

  p_heatmap1 = ggplot(dat, aes(x = Strain, y = Condition, fill = Base_model_aug_resid)) +
                geom_tile() +
                scale_fill_gradient2(low = ('cyan'), mid = 'black', high = 'yellow') +
                theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
                labs(fill = "Residual") +
                ggtitle("Expression ~ 1")

  p_heatmap2 = ggplot(dat, aes(x = Strain, y = Condition, fill = Strain_model_aug_resid)) +
                geom_tile() +
                scale_fill_gradient2(low = ('cyan'), mid = 'black', high = 'yellow') +
                theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
                labs(fill = "Residual") +
                ggtitle("Expression ~ Strain")

  p_heatmap3 = ggplot(dat, aes(x = Strain, y = Condition, fill = Condition_model_aug_resid)) +
                geom_tile() +
                scale_fill_gradient2(low = ('cyan'), mid = 'black', high = 'yellow') +
                theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
                labs(fill = "Residual") +
                ggtitle("Expression ~ Condition")

  p_heatmap4 = ggplot(dat, aes(x = Strain, y = Condition, fill = Full_model_aug_resid)) +
                geom_tile() +
                scale_fill_gradient2(low = ('cyan'), mid = 'black', high = 'yellow') +
                theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
                labs(fill = "Residual") +
                ggtitle("Expression ~ Strain + Condtion")



  grid.arrange(p_heatmap1, #+ theme(legend.position="none", axis.text.x=element_blank(), axis.text.y=element_blank()),
               p_heatmap2, #+ theme(legend.position="none", axis.text.x=element_blank(), axis.text.y=element_blank()),
               p_heatmap3, #+ theme(legend.position="none", axis.text.x=element_blank(), axis.text.y=element_blank()),
               p_heatmap4) #+ theme(legend.position="none", axis.text.x=element_blank(), axis.text.y=element_blank()))
}
residual_plot('ATP8', genes)
```


What makes the model week? are there any features of the genes that have low scores? $R^2$ is a general thing for all 300 data points, so its not like its pulling out massive outliers (FUS1 and HSP12 have very hight $R^2$)

Proving that this model is actually performing well:

What is the R^2 for this data vs data that is row shuffled, column shuffled, or completely shuffled?
