## The full linear model of Condition and Stress identity accurately describes gene expression data
(or at least for most genes)

1. What's happening: I'm going to build the linear model
2. I'm going to show that it has good performance over most genes. 

### Building the model:

```{r, messages = FALSE}
rm(list=ls())

load('../../input/images/normalized_data.RData')

library(tidyr)
library(purrr)
library(dplyr)
library(broom)
library(purrr)
library(tibble)
library(heatmap3)
library(ggplot2)
#library(magrittr)
#library(multidplyr)

#vlog = vlog - rowMeans(vlog)
#row_sd = apply(vlog,1,sd)
#vlog = apply(vlog,2,function(x) x / row_sd)

vlog = as.data.frame(vlog)
vlog$Gene = rownames(vlog)
measurements = gather(vlog, Sample_Name, Expression, -Gene)
#measurements %<>% as_tibble()

# should be a left join on the meta subsetted
meta_sub = as.data.frame(meta) %>% select(Sample_Name, Condition, Strain)

measurements = left_join(measurements, meta_sub)

lin_mod = function(formula) {
  function(data, ...){
    map(data, ~ lm(formula, data = .))
  }
}

list_model <- list(Strain_model = Expression ~ Strain,
                   Condition_model = Expression ~ Condition,
                   Full_model = Expression ~ Strain + Condition,
                   Base_model = Expression ~ 1) %>%
  lapply(lin_mod)

genes = measurements %>%
  group_by(Gene) %>%
  mutate(Gene_range = diff(range(Expression)),
         Gene_sd = sd(Expression),
         Gene_mean = mean(Expression),
         Norm_Expression = (Expression - Gene_mean) / Gene_sd) %>%
  filter(Gene_range > 1) %>%
  nest() %>%
  mutate_at(.vars=("data"),.funs=list_model)

```
So we've built a model for each gene. lets quickly look at the data frame 


```{r}
colnames(t2g)[1] = 'Gene'
genes = genes %>% left_join(t2g)
#knitr::kable(head(genes,100))
colnames(genes)
```

In order to asses the quality for each fit, lets calculate the $R^2$, or the percentage of variance explained for each gene from the model.

```{r}
mod_q = genes %>%
  select(Gene, name, data, ends_with('model')) %>%
  mutate_at(.vars=vars(ends_with('model')), 
            .funs = funs(glance = map(.,glance))) %>%
  select(Gene, name, ends_with('glance')) %>%
  gather(model_type, model_data, -Gene, -name) %>% 
  mutate(model_type = map(model_type, function(x) strsplit(x,'_')[[1]][1])) %>% 
  unnest()

head(mod_q)
```

Lets look at the $R^{2}$ for each gene in the full model:

```{r}
mod_q %>% 
    filter(model_type == 'Full') %>% 
    ggplot(aes(x=r.squared)) + 
    geom_density(fill='blue', alpha=.5) +
    ggtitle('Percentage of variance explained by the full model')
```
 Okay, so for most genes, the model does well. For what genes does it perform badly?
 
 
```{r}
mod_q %>% 
    filter(model_type == 'Full' & r.squared < .4) %>% 
    arrange(r.squared)
```
What makes the model week? are there any features of the genes that have low scores? $R^2$ is a general thing for all 300 data points, so its not like its pulling out massive outliers (FUS1 and HSP12 have very hight $R^2$)

Proving that this model is actually performing well:

What is the R^2 for this data vs data that is row shuffled, column shuffled, or completely shuffled?
